Simulations of gravity-driven flow of binary liquids in microchannels

### 2011-03-30 Text/paper.tex
Finished revisions. Waiting a reply from Jos.

\begin{abstract}
%This work shows that the free-energy binary liquid lattice Boltzmann method is able to
%reproduce the gravity driven microchannel flow of a gas bubble through liquid. In the regime of low
%capillary
%numbers
%$\Ca$ (defined in Eq.~\ref{capillary:number:definition}), the flow is thoroughly described by
%Bretherton \cite{bretherton},
%with the deposition film thickness proportional to $\Ca^{2/3}$. However, there are deviations
%from the Bretherton correlation for $\Ca\geq 0.005$, studied experimentally and numerically. This
%work aims to examine the applicability of the lattice Boltzmann binary liquid model to simulations
%of bubble flow in liquids for a range of $\Ca$ numbers. The binary
%liquid model is a continuous-interface multiphase model allowing to properly resolve the interface.
%It was found that the interface thickness needs to be $50$
%percent of the film thickness for the simulation to give convergent results. When
%the convergence criterion is fulfilled, the results are in a good agreement with the
%numerical simulations of \citet{giavedoni-numerical} in the range of $\Ca$ from $0.01$ to $1$. 
In this work a free-energy binary liquid lattice-Boltzmann scheme is used to simulate
Taylor/Bretherton flow in a micro-channel where elongated gas bubbles move through a liquid with
thin liquid films between the bubbles and the channel walls. The numerical scheme has a diffuse
interface, and a main focus of our work is to assess resolution requirements for correctly resolving
the liquid film and bubble motion. The simulations are two-dimensional and span a capillary number
range of $0.05$ to $1.0$ where the capillary number is based on the liquid’s dynamic viscosity, the
velocity of the bubble, and the interfacial tension. The flow is driven by a body force, and
periodic boundary conditions apply in the streamwise direction. We obtain grid independent results
as long as the liquid film thickness is at least two times the width of the diffuse interface, with
film thicknesses in accordance to literature results. We also show that the results in terms of film
thicknesses are largely insensitive to the liquid-gas viscosity ratio and wettability parameters. 
\end{abstract}

@@

\section{Introduction}
The Taylor/Bretherton \cite{bretherton} flow deals with long gas bubbles moving through liquid in
narrow channels. It was found that the front meniscus film thickness is proportional
to $\Ca^{2/3}$ for capillary numbers smaller than $0.003$, where the capillary number is defined as:
\begin{equation}
\label{capillary:number:definition}
\Ca=\frac{\mu_\mathrm{liq} U_\mathrm{bubble}}{\gamma},
\end{equation}
where $\mu_{liq}$ is the liquid viscosity, $U_{bubble}$ is the bubble velocity, and
$\gamma$ is the interfacial tension between gas and liquid.

While there are many works which simulate the Bretherton problem by the boundary value approach
\cite{ingham-plates,heil-bretherton}, those methods are of limited applicability for problems
involving complex geometries, free interface motion, or for problems involving coalescence and/or
droplet breakup. The continuous interface models are more flexible for those kind of simulations.
However, if the interface is smeared out over several grid nodes, the question of
proper film resolution in comparison with the interface resolution arises.  This work is focused on
exploring the parameter range of the binary liquid lattice
Boltzmann method for correctly resolving the flow in a range of capillary numbers.

The lattice Boltzmann method has emerged as a successful method to simulate
a wide variety of phenomena including hydrodynamics \cite{yu}, thermal flows
\cite{karlin-minimalmodels}, microflows \cite{ansumali-small-knudsen},
ferrofluids \cite{kuzmin-aniso}, and multiphase flows
\cite{swift,Shan-chen:extended}. Thanks to its kinetic nature, LBM as a particle
method easily tackles complex geometries and allows incorporation of
physical phenomena on the microscopic level, as in the case of multiphase models. Most
multiphase lattice Boltzmann models \cite{swift, Shan-chen:extended} resolve
the interface using continuous interface methods where the interface spans over several grid nodes.
Such a representation
brings issues of the film thickness resolution versus interface
resolution -- the diffuse interface should be dealt with in such a way as to have a
negligible effect on the physics of the film.

The binary liquid free-energy LB model due to \citet{swift} we used
simulates two liquids with the assumption of uniform overall
density. While the classical Bretherton problem is stated for gas and liquid, which are of
significantly
different densities and viscosities, it was indicated \cite{bretherton} that inertia effects were
neglected. Moreover, the results of \citet{giavedoni-numerical} and \citet{heil-bretherton} show
{\color{red} insignificant} Reynolds number effects on the film thickness for a relatively wide
range of Reynolds
numbers. 
{\color{red} For example, \citet{giavedoni-numerical} suggested that the Reynolds number effects
are negligible for $\Ca\leq0.05$ and have moderate changes for $\Ca>0.05$ in the range of Reynolds
number from $0$ to $70$. Later on, citet{heil-bretherton} extended results upto Reynolds number
$300$. The indicate that while the influence of the established film thickness is insignificant
($7$ percent from the film thickness measured at $Re=0$). However, the Reynolds number influence
significantly the pressure distrubution and the flow field near the front bubble tip. In the present
simulations the largest Reynolds number we achieve is less than $20$ and it allows to neglect the
inertia effects.} 
Therefore, the major governing factor for microchannel flows is not the density ratio, but the
viscosity ratio. Moreover, the results based on the uniform density, as in the case of the LBM
binary liquid model, are in good agreement with other simulations
\cite{giavedoni-numerical,heil-bretherton} in which inertia effects were taken into account.

The goal of this paper is to understand how to choose the
binary liquid model parameters, the viscosity ratio and the grid
resolution to obtain correct flow physics with an emphasis on film thicknesses.

@@

\section{Bretherton bubble flow}
Bretherton \cite{bretherton} studied a long bubble
moving in a tube filled with liquid. It was found that the deposition film thickness
is proportional to $\Ca^{2/3}$ in the range of small capillary numbers. Later,
it was realized \cite{wong-films,wong-pressure} that the film thickness
is proportional to $\Ca^{2/3}$ only in the certain region behind the front meniscus and
that the film thickness
varies over the bubble length for bubbles of finite length. Numerical
simulations \cite{giavedoni-numerical} and experimental studies
\cite{kreutzer-pressure-drop} showed a deviation from the $\Ca^{2/3}$ rule for capillary numbers
larger than $0.003$.
To consistently  predict a flow pattern for capillaries in
different ranges of parameters, simulations have been conducted and have been typically validated with the
small capillary numbers Bretherton problem.

There are a number of numerical methods which
were used for the simulation of the Taylor/Bretherton flow.
\citet{vanbaten-circular} studied the mass transfer and film
thickness for rising bubbles in a circular capillary using the finite volume method.
\citet{kreutzer-pressure-drop} also used the finite volume method to perform
simulations of a circular capillary for a number of different
Reynolds and capillary numbers. \citeauthor{wong-films} \cite{wong-films,wong-pressure} studied
three-dimensional bubbles in
polygonal capillaries and calculated bubble shapes in the
slug cross sections and menisci appearance.
\citet{heil-bretherton} and \citet{ingham-plates} studied gas finger propagation in
a two-dimensional channel for a range of Reynolds and capillary
numbers using the finite element method. \citet{giavedoni-numerical} performed cross validation of the
finite element solution with previously published results.
The solutions were obtained for circular and planar cases.

@@

One should acknowledge the works of \citet{pagonabarraga-fingers} on menisci
in thin films for fingering phenomena. \citet{sehgal-microchannel} performed lattice Boltzmann
simulations of two-dimensional channel flows for a relatively large capillary numbers, and
found discrepancies with the classical Bretherton theory, which
is limited to the low capillary number regime \cite{giavedoni-numerical}. Authors didn't take into
the account recent studies extending microchannel simulations for the capillary numbers
beyond the Bretherton regime. Thus, limiting comparison with other established CFD methods. Author
used the Shan-Chen model \cite{Shan-chen:extended} to simulate the multiphase behavior. It is also
the continuous interface model, but the study of grid dependence is absent.

@@

Resolving the interface on a fine level for capillary numbers smaller than
$0.003$ is computationally expensive even in the two-dimensional case
(the amount of memory necessary to perform the simulation would be of the order
of tens or hundreds gigabytes).  This work is therefore based on a
comparison of the results obtained with the known data
validated for the Bretherton bubble flow.  In this paper, we present
techniques for initialization of the simulations and discuss optimal
parameter ranges for the binary liquid LB model.

\section{Lattice Boltzmann binary liquid model}
The lattice Boltzmann equation (LBE) operates on a rectangular grid representing the
physical domain. It utilizes
probability distribution functions (also known as particle populations)
containing information about
macroscopic variables, such as fluid density and momentum. LBE consists of
two parts: a local collision step, and a propagation step which transports
information from one node to another in certain
directions specified by the discrete velocity set.
The LBE is typically implemented as follows:
\begin{equation}
\label{standard:implementation}
\begin{aligned}
&f_i^{*}(\bm{x},t)=\omega f_i^{eq}(\bm{x},t)-(1-\omega) f_i(\bm{x},t) +
F_i,&&\text{ collision step}\\
&f_i(\bm{x}+\bm{c_i},t+1)=f_i^{*}(\bm{x},t),&&\text{ propagation step}, 
\end{aligned}
\end{equation}
where $f_i$ is the probability distribution function in the direction $\bm{c_i}$,
{\color{red} $f_i^{q}$ is the equilibrium probability distribution function}, $\omega$ is the
relaxation parameter, and $F_i$ is the external force population.{\color{red} The force population
is connected with the force and is implemented in the current work due to work \cite{guo}.}

@@

The model we use is a two-dimensional nine-velocity (D2Q9) model,
with equilibrium populations \cite{pooley-contact}:
\begin{equation}
\label{set:equilibrium:binary}
\begin{aligned}
&f_i^{eq}&&=w_i 
\biggl(3
p_0 - k \phi \Delta \phi
+\frac{u_{\alpha}c_{i\alpha}}{c_s^2}+\frac{Q_{i\alpha\beta}u_{\alpha } u_ {
\beta}}{2 c_s^4}\biggr)\\
&&&+k\bigl(w_i^{xx} (\partial_x \phi)^2+w_i^{yy} (\partial_y \phi)^2 +w_i^{xy} \partial_x
\phi \partial_y \phi \bigr), 1\leq i \leq 8\\
&f_0^{eq}&&=\rho-\sum_{i\neq0}{f_i^{eq}}\\
&g_i^{eq}&&=w_i\left(\Gamma \mu + \frac{\phi c_{i\alpha} u_{i\alpha}}{c_s^2}+\phi
\frac{Q_{i\alpha\beta}u_{\alpha}u_{\beta}}{2 c_s^4}\right), 1\leq i \leq 8 \\
&g_0^{eq}&&=\phi-\sum_{i\neq0}{g_i^{eq}}\quad,
\end{aligned}
\end{equation}
where $\Gamma$ is the mobility parameter; the chemical potential
$\mu=-A\phi+A\phi^3-k\Delta\phi$; $k$ is the parameter related to the surface
tension; $A$ is the parameter of the free-energy model. The bulk pressure
is expressed as $p_0=c_s^2 \rho +A (-0.5 \phi^2+0.75 \phi^4)$ {\color{red} with
the sound speed $c_s^2=1/3$}. 
Parameters specific to the D2Q9 grid are the weights
$w_i=\left\{\frac{4}{9},\frac{1}{9},\frac{1}{9},\frac{1}{9},\frac{1}{9},
\frac{1}{36},\frac{1}{36},\frac{1}{36},\frac{1}{36}\right\}$, and the tensor
$Q_{i\alpha\beta}=c_{i\alpha} c_{i\beta} - c_s^2 \delta_{\alpha\beta}$.  
{\color{red}
Other weights 
%related to the
%inclusion of the surface tension coefficient into the equations 
are as follows:}
$w^{xx}_{1-2}=w^{yy}_{3-4}=1/3$, $w^{xx}_{3-4}=w^{yy}_{1-2}=-1/6$,
$w^{xx}_{5-8}=w^{yy}_{5-8}=-1/24$, $w^{xy}_{1-4}=0$, $w^{xy}_{5-6}=1/4$ and
$w^{xy}_{7-8}=-1/4$. The set of equations (\ref{set:equilibrium:binary}) restores the
macroscopic
fluid equations as:
\begin{equation}
\begin{aligned}
&\partial_t \rho+ \partial_{\alpha} \rho u_{\alpha}=0\\
&\rho\left(\partial_t+u_{\beta}\partial_{\beta}\right) u_{\alpha}= F_{\alpha}
-\partial_{\beta}P_{\alpha \beta} +
\nu\partial_{\beta}\left(\partial_{\alpha}u_{\beta}+\partial_{\beta} u_{\alpha}\right)\\
&\partial_t \phi + \partial_{\alpha} \phi u_{\alpha}=M \partial^2_{\beta\beta} \mu,
\end{aligned}
\label{binary:fluid:system}
\end{equation}
where $\nu=c_s^2 (\tau-1/2)$ is the viscosity,
$M=\Gamma(\tau_{\phi}-1/2)$ is the mobility parameter, and $\tau=\frac{1}{\omega}$ and $\tau_{\phi}$
are the relaxation parameters of density and phase fields, {\color{red}
$P_{\alpha\beta}=\Bigl(p_0-k\phi \Delta \phi -\frac{k}{2}|\nabla \phi|^2\Bigr)\delta_{\alpha\beta}
+ k \partial_{\alpha} \phi \partial_{\beta} \phi$  \cite{pooley-contact}.The interface tension value
in the framework of the binary liquid model is $\gamma=\sqrt{\frac{8 k
A}{9}}$. The inclusion of the interface tension to the momentum flux tensor is done through the
coefficients $k$, $A$ and weights $w_i^{\alpha\beta}$.

@@

\section{Results}
\subsection{The nondimensionalization and initialization procedure}
\label{section:init}
We now introduce the procedure of nondimensionalization, as well as
the initialization technique we used to determine the parameters necessary for
the simulation. The capillary number defined in Eq.~\ref{capillary:number:definition} governs the
interface thickness.
%The bubble velocity $U_{\mathrm{bubble}}$ is taken in the center of the bubble where the film
%thickness is examined.
While parameters such as $\mu_{\mathrm{liq}}$, $\gamma$ and the capillary number $\Ca$ are given
or defined through the model parameters, the bubble velocity $U_{\mathrm{bubble}}$ is not given
explicitly.
In our simulations we approximate $U_{bubble}$ with the
Poiseuille profile center velocity, which is given as:
\begin{equation}
\begin{aligned}
U_{\mathrm{bubble}} \approx \frac{H_{\mathrm{eff}}^2}{8
\mu_{\mathrm{liq}}}\frac{\mathrm{d}P}{\mathrm{d}x}\\
U_{\mathrm{bubble}}\approx
\frac{{(N_y-2)}^2}{\frac{8}{6}(\tau_{\mathrm{liq}}-\frac{1}{2})}\frac{\mathrm{d}P}{\mathrm{d}
x } ,
\end{aligned}
\end{equation}
where $N_y$ is the grid number in the vertical direction. As will be discussed further, the
effective channel width is defined as $H_{\mathrm{eff}}=N_y-2$ because of bounce-back nodes to
mimic no-slip conditions at the walls.
For the Poiseuille profile one can estimate the body force (pressure gradient) through the
capillary number and grid resolution:
\begin{equation}
\label{poiseuille:velocity:center}
\begin{aligned}
&\frac{\mathrm{d}P}{\mathrm{d}x}=\frac{8}{{(N_y-2)}^2}\gamma \Ca.
\end{aligned}
\end{equation}
In simulations the body force (\ref{poiseuille:velocity:center}) causes the bubble
velocity to differ from that implied by the Poiseuille relation because of the bubble presence in a
flow. That is
why an iteration procedure is needed to achieve the 
desired velocity and 
capillary number. 
However, the initialization body force given by (\ref{poiseuille:velocity:center}) is a good
starting approximation.

@@

{\color{red}
For demonstration purposes we present the simulation parameters choice for $\Ca=0.05$.
In these
examples we use $k=0.04$ and $A=0.04$ as optimal binary liquid model parameters.
That explicitly implies the surface tension to be $\sqrt{\frac{8 k
A}{9}}=0.0377$ and the characteristic length of the interface
$5\sqrt{\frac{k}{A}}=5$ lattice Boltzmann units (grid nodes). For the sake of simplicity and
stability, the relaxation times for liquid and gas phases are taken as
$\tau_{\mathrm{liq}}=2.5$ and $\tau_{\mathrm{gas}}=0.7$, which correspond
to the liquid over gas kinematic viscosity ratio equal to $10$.

@@

, and for Reynolds number $40$ equals $0.06$.  
\item[Grid choice] After pecifying the film thickness, one needs to choose the associate

@@

  resolve $6\%$ of the whole channel width with $12$ lattice units.  Thus, the channel

@@

each other. Thus, the whole grid size is $200 \times 3000$.

@@

necessary, starting with the Poiseuille pressure gradient force as the initial condition. We extracted

>>>>

\begin{abstract}
%This work shows that the free-energy binary liquid lattice Boltzmann method is able to
%reproduce the gravity driven microchannel flow of a gas bubble through liquid. In the regime of low
%capillary
%numbers
%$\Ca$ (defined in Eq.~\ref{capillary:number:definition}), the flow is thoroughly described by
%Bretherton \cite{bretherton},
%with the deposition film thickness proportional to $\Ca^{2/3}$. However, there are deviations
%from the Bretherton correlation for $\Ca\geq 0.005$, studied experimentally and numerically. This
%work aims to examine the applicability of the lattice Boltzmann binary liquid model to simulations
%of bubble flow in liquids for a range of $\Ca$ numbers. The binary
%liquid model is a continuous-interface multiphase model allowing to properly resolve the interface.
%It was found that the interface thickness needs to be $50$
%percent of the film thickness for the simulation to give convergent results. When
%the convergence criterion is fulfilled, the results are in a good agreement with the
%numerical simulations of \citet{giavedoni-numerical} in the range of $\Ca$ from $0.01$ to $1$. 
In this work a free-energy binary liquid lattice-Boltzmann scheme is used to simulate
Taylor/Bretherton flow in a micro-channel where elongated gas bubbles move through a liquid with
thin liquid films between the bubbles and the channel walls. The numerical scheme has a diffuse
interface, and a main focus of our work is to assess resolution requirements for correctly resolving
the liquid film and bubble motion. The simulations are two-dimensional and span a capillary number
range of $0.05$ to $1.0$ where the capillary number is based on the liquid dynamic viscosity, the
velocity of the bubble, and the interfacial tension. The flow is driven by a body force, and
periodic boundary conditions apply in the streamwise direction. We obtain grid independent results
as long as the liquid film thickness is at least twice the width of the diffuse interface, with
film thicknesses in accordance to literature results. We also show that the results in terms of film
thicknesses are largely insensitive to the liquid-gas viscosity ratio and wettability parameters. 
\end{abstract}

@@

\section{Introduction}
The Taylor/Bretherton \cite{bretherton} flow deals with long gas bubbles moving through liquid in
narrow channels, Fig. \ref{fig:classical:bretherton}. It was found that the front meniscus film
thickness is proportional
to $\Ca^{2/3}$ for capillary numbers smaller than $0.003$, where the capillary number is defined as:
\begin{equation}
\label{capillary:number:definition}
\Ca=\frac{\mu_\mathrm{liq} U_\mathrm{bubble}}{\gamma},
\end{equation}
where $\mu_{liq}$ is the liquid viscosity, $U_{bubble}$ is the bubble velocity, and
$\gamma$ is the interfacial tension between gas and liquid.
\begin{figure}
\includegraphics[width=\textwidth]{Figures/benchmark_bretherton.eps}
\caption{The classical Bretherton problem layout. The gas bubble
propagates with the steady velocity $U_{\mathrm{bubble}}$ through the liquid media and deposits
the film, where thickness to the height of the channel $H_{\mathrm{eff}}$ ratio is 
$\delta$.\label{fig:classical:bretherton}} 
\end{figure}

While there are many works which simulate the Bretherton problem by the boundary value approach
\cite{ingham-plates,heil-bretherton}, those methods are of limited applicability for problems
involving complex geometries, free interface motion, or for problems involving coalescence and/or
droplet breakup. The continuous interface models are more flexible for those kinds of simulations.
However, if the interface is smeared out over several grid nodes, the question of
proper film resolution in comparison with the interface resolution arises.  This work is focused on
exploring the parameter range of the binary liquid lattice
Boltzmann method to resolve correctly the flow in a range of capillary numbers.

The lattice Boltzmann method (LBM) has emerged as a successful method to simulate
a wide variety of phenomena including hydrodynamics \cite{yu}, thermal flows
\cite{karlin-minimalmodels}, microflows \cite{ansumali-small-knudsen},
ferrofluids \cite{kuzmin-aniso}, and multiphase flows
\cite{swift,Shan-chen:extended}. Thanks to its kinetic nature, LBM as a particle
method easily tackles complex geometries and allows incorporation of
physical phenomena on the microscopic level, as in the case of multiphase models. Most
multiphase lattice Boltzmann models \cite{swift, Shan-chen:extended} resolve
the interface using continuous interface methods where the interface spans over several grid nodes.
Such a representation
brings issues of the film thickness resolution versus interface
resolution -- the diffuse interface should be dealt with in such a way as to have a
negligible effect on the physics of the film.

The binary liquid free-energy LB model due to \citet{swift} we used
simulates two liquids with the assumption of uniform overall
density. While the classical Bretherton problem is stated for gas and liquid, which are of
significantly
different densities and viscosities, it was indicated \cite{bretherton} that inertia effects were
neglected. Moreover, the results of \citet{giavedoni-numerical} and \citet{heil-bretherton} show
{\color{red} insignificant} Reynolds number effects on the film thickness for a relatively wide
range of Reynolds
numbers. 
{\color{red} For example, \citet{giavedoni-numerical} suggested that the Reynolds number effects
are negligible for $\Ca\leq0.05$ and have moderate changes for $\Ca>0.05$ in the range of Reynolds
number from $0$ to $70$. Later on, citet{heil-bretherton} extended results upto Reynolds number
$300$. The indicate that while the influence of the established film thickness is insignificant
($7$ percent from the film thickness measured at $Re=0$). However, the Reynolds number influence
significantly the pressure distrubution and the flow field near the front bubble tip. In the present
simulations the largest Reynolds number we achieve is less than $20$ and it allows to neglect the
inertia effects.} 
Therefore, the major governing parameter for microchannel Bretherton flows is not the density ratio,
but the
viscosity ratio. Moreover, the results based on the uniform density, as in the case of the LBM
binary liquid model, are in good agreement with other simulations
\cite{giavedoni-numerical,heil-bretherton} in which inertia effects were taken into account.

Resolving the interface on a fine level for capillary numbers smaller than
$0.003$ is computationally expensive even in the two-dimensional case
(the amount of memory necessary to perform the simulation would be of the order
of tens or hundreds gigabytes).  This work is therefore based on a
comparison of the results obtained with the known data
validated for the Bretherton bubble flow.  In this paper, we present
techniques for initialization of the simulations and discuss optimal
parameter ranges for the binary liquid LB model, as the binary liquid parameters, the viscosity
ratio and the grid resolution. The emphasis is to obtain correct flow physics, especially in terms
of film thicknesses.

%The goal of this paper is to understand how to choose the
%binary liquid model parameters, the viscosity ratio and the grid
%resolution to obtain correct flow physics with an emphasis on film thicknesses.

@@

\section{Bretherton bubble flow}
Bretherton \cite{bretherton} studied a long bubble
moving in a tube filled with liquid. It was found that the film thickness
is proportional to $\Ca^{2/3}$ in the range of small capillary numbers. Later,
it was realized \cite{wong-films,wong-pressure} that the film thickness
is proportional to $\Ca^{2/3}$ only in the certain region behind the front meniscus and
that the film thickness
varies over the bubble length for bubbles of finite length. Numerical
simulations \cite{giavedoni-numerical} and experimental studies
\cite{kreutzer-pressure-drop} showed a deviation from the $\Ca^{2/3}$ rule for capillary numbers
larger than $0.003$.
To consistently  predict a flow pattern for capillaries in
different ranges of parameters, simulations have been conducted and have been typically validated with the
small capillary numbers Bretherton problem.

There are a number of numerical methods which
were used for the simulation of the Taylor/Bretherton flow.
\citet{vanbaten-circular} studied the mass transfer and film
thickness for rising bubbles in a circular capillary using the finite volume method.
\citet{kreutzer-pressure-drop} also used the finite volume method to perform
simulations of a circular capillary for a number of different
Reynolds and capillary numbers. \citeauthor{wong-films} \cite{wong-films,wong-pressure} studied
three-dimensional bubbles in
polygonal capillaries and calculated bubble shapes in the
slug cross sections and menisci appearance.
\citet{heil-bretherton} and \citet{ingham-plates} studied gas finger propagation in
a two-dimensional channel for a range of Reynolds and capillary
numbers using the finite element method. \citet{giavedoni-numerical} performed cross validation of the
finite element solution with previously published results.
The solutions were obtained for circular and planar geometries.

@@

One should acknowledge the work of \citet{pagonabarraga-fingers} on menisci
in thin films for fingering phenomena. \citet{sehgal-microchannel} performed lattice Boltzmann
simulations of two-dimensional channel flows for relatively large capillary numbers, and
found discrepancies with the classical Bretherton theory, which
is limited to the low capillary number regime \cite{giavedoni-numerical}. Authors didn't take into
the account recent studies extending microchannel simulations for the capillary numbers
beyond the Bretherton regime. Thus, limiting comparison with other established CFD methods. Author
used the Shan-Chen model \cite{Shan-chen:extended} to simulate the multiphase behavior. It is also
the continuous interface model, but the study of grid dependence is absent.

@@

\section{Lattice Boltzmann binary liquid model}
The lattice Boltzmann equation (LBE) operates on a rectangular grid representing the
physical domain. It utilizes
probability distribution functions (also known as particle populations)
containing information about
macroscopic variables, such as fluid density and momentum. LBE consists of
two parts: a local collision step, and a propagation step which transports
information from one node to another along some 
directions specified by the discrete velocity set.
The LBE is typically implemented as follows:
\begin{equation}
\label{standard:implementation}
\begin{aligned}
&f_i^{*}(\bm{x},t)=\omega f_i^{eq}(\bm{x},t)-(1-\omega) f_i(\bm{x},t) +
F_i,&&\text{ collision step}\\
&f_i(\bm{x}+\bm{c_i},t+1)=f_i^{*}(\bm{x},t),&&\text{ propagation step}, 
\end{aligned}
\end{equation}
where $f_i$ is the probability distribution function in the direction $\bm{c_i}$,
{\color{red} $f_i^{q}$ is the equilibrium probability distribution function}, $\omega$ is the
relaxation parameter, and $F_i$ is the external force population.{\color{red} The force population
is connected with the force and is implemented in the current work due to work \cite{guo}.}

@@

The model we use is a two-dimensional nine-velocity (D2Q9) model,
with equilibrium populations \cite{pooley-contact}:
\begin{equation}
\label{set:equilibrium:binary}
\begin{aligned}
&f_i^{eq}&&=w_i 
\biggl(3
p_0 - k \phi \Delta \phi
+\rho\frac{u_{\alpha}c_{i\alpha}}{c_s^2}+\rho \frac{Q_{i\alpha\beta}u_{\alpha } u_ {
\beta}}{2 c_s^4}\biggr)\\
&&&+k\bigl(w_i^{xx} (\partial_x \phi)^2+w_i^{yy} (\partial_y \phi)^2 +w_i^{xy} \partial_x
\phi \partial_y \phi \bigr), 1\leq i \leq 8\\
&f_0^{eq}&&=\rho-\sum_{i\neq0}{f_i^{eq}}\\
&g_i^{eq}&&=w_i\left(\Gamma \mu + \phi\frac{ c_{i\alpha} u_{i\alpha}}{c_s^2}+\phi
\frac{Q_{i\alpha\beta}u_{\alpha}u_{\beta}}{2 c_s^4}\right), 1\leq i \leq 8 \\
&g_0^{eq}&&=\phi-\sum_{i\neq0}{g_i^{eq}}\quad,
\end{aligned}
\end{equation}
where $\Gamma$ is the mobility parameter; the chemical potential
$\mu=-A\phi+A\phi^3-k\Delta\phi$; $k$ is the parameter related to the surface
tension; $A$ is the parameter of the free-energy model. The bulk pressure
is expressed as $p_0=c_s^2 \rho +A (-0.5 \phi^2+0.75 \phi^4)$ {\color{red} with
the sound speed $c_s^2=1/3$}. 
Parameters specific to the D2Q9 grid are the weights
$w_i=\left\{\frac{4}{9},\frac{1}{9},\frac{1}{9},\frac{1}{9},\frac{1}{9},
\frac{1}{36},\frac{1}{36},\frac{1}{36},\frac{1}{36}\right\}$, and the tensor
$Q_{i\alpha\beta}=c_{i\alpha} c_{i\beta} - c_s^2 \delta_{\alpha\beta}$.  
{\color{red}
Other weights 
%related to the
%inclusion of the surface tension coefficient into the equations 
are as follows:}
$w^{xx}_{1-2}=w^{yy}_{3-4}=1/3$, $w^{xx}_{3-4}=w^{yy}_{1-2}=-1/6$,
$w^{xx}_{5-8}=w^{yy}_{5-8}=-1/24$, $w^{xy}_{1-4}=0$, $w^{xy}_{5-6}=1/4$ and
$w^{xy}_{7-8}=-1/4$. The set of equations (\ref{set:equilibrium:binary}) restores the
macroscopic
fluid equations as:
\begin{equation}
\begin{aligned}
&\partial_t \rho+ \partial_{\alpha} \rho u_{\alpha}=0\\
&\rho\left(\partial_t+u_{\beta}\partial_{\beta}\right) u_{\alpha}= F_{\alpha}
-\partial_{\beta}P_{\alpha \beta} +
\nu\partial_{\beta}\left(\partial_{\alpha}u_{\beta}+\partial_{\beta} u_{\alpha}\right)\\
&\partial_t \phi + \partial_{\alpha} \phi u_{\alpha}=M \partial^2_{\beta\beta} \mu,
\end{aligned}
\label{binary:fluid:system}
\end{equation}
where $\nu=c_s^2 (\tau-1/2)$ is the viscosity,
$M=\Gamma(\tau_{\phi}-1/2)$ is the mobility parameter, and $\tau=\frac{1}{\omega}$ and $\tau_{\phi}$
are the relaxation parameters of density and phase fields, {\color{red}
$P_{\alpha\beta}=\Bigl(p_0-k\phi \Delta \phi -\frac{k}{2}|\nabla \phi|^2\Bigr)\delta_{\alpha\beta}
+ k \partial_{\alpha} \phi \partial_{\beta} \phi$  \cite{pooley-contact}.The interface tension value
in the framework of the binary liquid model is $\gamma=\sqrt{\frac{8 k
A}{9}}$. The inclusion of the interface tension to the momentum flux tensor is done through the
coefficients $k$, $A$ and weights $w_i^{\alpha\beta}$.

@@

\section{Results}
\subsection{The nondimensionalization and initialization procedure}
\label{section:init}
We now introduce the procedure of nondimensionalization, as well as
the initialization technique we used to determine the parameters necessary for
the simulation. The capillary number defined in Eq.~\ref{capillary:number:definition} governs the
interface thickness.
%The bubble velocity $U_{\mathrm{bubble}}$ is taken in the center of the bubble where the film
%thickness is examined.
{\color{red} Parameters as $\mu_{\mathrm{liq}}$, $\gamma$ are usually taken in the already defined
stable range. The capillary number $\Ca$ is supplied in order to simulate.} 
In comparison, the bubble
velocity $U_{\mathrm{bubble}}$ is not given
explicitly.
In our simulations we approximate $U_{\mathrm{bubble}}$ with the
Poiseuille profile center velocity, which is given as:
\begin{equation}
\begin{aligned}
&U_{\mathrm{bubble}} \approx \frac{H_{\mathrm{eff}}^2}{8
\mu_{\mathrm{liq}}}\frac{\mathrm{d}P}{\mathrm{d}x}\\
&U_{\mathrm{bubble}}\approx
\frac{{(N_y-2)}^2}{\frac{8}{6}(\tau_{\mathrm{liq}}-\frac{1}{2})}\frac{\mathrm{d}P}{\mathrm{d}
x } ,
\end{aligned}
\end{equation}
where $N_y$ is the number of nodes in the vertical direction. As will be discussed further, the
effective channel width is defined as $H_{\mathrm{eff}}=N_y-2$ because of bounce-back nodes to
mimic no-slip conditions at the walls.
For the Poiseuille profile one can estimate the body force (pressure gradient) through the
capillary number and grid resolution:
\begin{equation}
\label{poiseuille:velocity:center}
\begin{aligned}
&\frac{\mathrm{d}P}{\mathrm{d}x}=\frac{8}{{(N_y-2)}^2}\gamma \Ca.
\end{aligned}
\end{equation}
In simulations the body force (\ref{poiseuille:velocity:center}) causes the bubble
velocity to differ from that implied by the Poiseuille relation because of the bubble presence in a
flow. That is
why an iteration procedure is needed to achieve the 
desired velocity and 
capillary number. 
However, the initialization body force given by (\ref{poiseuille:velocity:center}) is a good
starting approximation.

@@

{\color{red}
For demonstration purposes we present the simulation parameters choice for $\Ca=0.05$.
In these
examples we use $k=0.04$ and $A=0.04$ as optimal binary liquid model parameters.
That explicitly implies the surface tension to be $\sqrt{\frac{8 k
A}{9}}=0.0377$ and the characteristic width of the interface
$5\sqrt{\frac{k}{A}}=5$ lattice Boltzmann units (grid nodes). For the sake of simplicity and
stability, the relaxation times for liquid and gas phases are taken as
$\tau_{\mathrm{liq}}=2.5$ and $\tau_{\mathrm{gas}}=0.7$, which correspond
to the liquid over gas kinematic viscosity ratio equal to $10$.

@@

, and for Reynolds number $40$ equals $0.06-0.07$.  
\item[Grid choice] After specifying the film thickness, one needs to choose the associate

@@

  resolve $6\%$ of the whole channel width with $12$ lattice units.  Thus, the physical channel

@@

each other. Thus, the whole physical grid size is $200 \times 3000$. Note that in the case of the
half-way bounce-back walls \cite{yu} which are used in the
simulations one needs to calculate the film thickness as:
\delta=(\phi_0-0.5)/(N_y-2),
where $\phi_0$ is the grid coordinate where phase field is $0$, $N_y-2$
is the effective channel height.
If the grid size in the $y$ direction is $N_y$, then one has $N_y-1$ regions between the grid nodes
which represent
the physical domain. The effective wall location is in the middle between bounce-back and fluid
nodes giving
overall $N_y-2$ nodes representing the fluid. Note that it is a simplification to impose the
boundary in the middle between the bounce-back
node and the fluid node. The location of the wall is
viscosity dependent \cite{ginzburg-multireflection}.  The effective location of the
wall for the
multiphase models to the best authors' knowledge is not yet derived. Thus, the simulation grid size
is $202\time 3000$. 

@@

necessary, starting with the Poiseuille pressure gradient force as the initial condition. 
{\color{red} Note that in the present work the shooting method is not used but it is presented as
a possible remedy to exactly match the required capillary number.} 
We
extracted
